---
title: "Analyzing Moisturizer Ingredient Lists"
output: html_document
---

```{r, echo=FALSE}
knitr::opts_chunk$set(global.par = TRUE)
library(here)
```

### Introduction

I'm pretty sure skincare companies are just making stuff up. They aren't doing research. They aren't inventing anything. They are only focused on getting people to buy their products. So they put all of their effort and money into marketing and copying the things that other companies have successfully marketed.

I wanted to get a sense of what's actually in these products, and I think just reading the ingredient lists makes it hard to tell what is going on. I know enough chemistry to understand what the ingredient names mean. I know data analysis. So let's take a more global approach to understanding skincare products, and try to analyze the data. 


### Getting the data

First, we have to get some data. I got the best-seller results from [Target](https://www.target.com/) and [Sephora](https://www.sephora.com/), and created a [google sheet](https://docs.google.com/spreadsheets/d/e/2PACX-1vS4wAroj1XvB4aiuOg97IdS6xNd2x0wsylwXOYZZlCAysx3n5SA9Fi3IcLrbWGe_e7cuhzW2_fOl8qj/pubhtml) with each product's brand name, product name, price, size, and product link on [INCIDecoder](https://www.incidecoder.com), which is where we will be scraping the ingredient lists. Sadly, the Target and Sephora websites are difficult to scrape. I had to gather all of the data by hand :( I might try to implement that in the future with [RSelenium](https://docs.ropensci.org/RSelenium/). 

The google sheet is imported using the [googlesheets4 R package](https://googlesheets4.tidyverse.org/).
```{r message=FALSE}
library(googlesheets4)
productSheet <- read_sheet("1LVIoYIZDoDtc26z5JMHpv_7a34aqSZwdFsazG8_JlfY")
```

Here are the top few entries:
```{r echo=FALSE}
kable(head(productSheet, n = 5L))
```

Next, we loop through all of the INCIDecoder product URLs and grab the ingredient lists using the [rvest package](https://rvest.tidyverse.org/).
```{r}
# library(rvest)
# ingredientLists <- list()
# for(i in productSheet$INCIDECODER_URL){
#   productUrl <- paste0("https://incidecoder.com/products/", i)
#   productHtml <- read_html(productUrl)
#   productData <- html_table(productHtml)[[1]]$`Ingredient name`
#   ingredientLists[[i]] <- productData
# }
ingredientLists <- readRDS(paste0(here(), "/data/ingredientLists"))
```

Here are the the first 5 ingredients from the first 3 products:
```{r echo=FALSE}
kable(sapply(ingredientLists[1:3], head, 5))
```

As you might notice. We have some problems with how ingredients are labeled. The main ingredient in each of those products is water. Of course, these companies can't just call it water. They have to be special and call it "Aqua," or "Eau," or "Purified." And don't get me started on "La Roche-Posay Prebiotic Thermal Water."

### Cleaning up ingredient names

How many names for water can there be?
```{r echo=FALSE, fig.width=10, fig.height=7}
ingredientsAll <- as.vector(do.call(c, ingredientLists, quote=T))
waterNames <- ingredientsAll[grep("Water|Aqua|Eau", ingredientsAll)]
waterNames <- gsub("[^ -~]+", "", waterNames)
waterNames <- waterNames[-grep("Citrullus|Rosa|Vitis", waterNames)]
waterCounts <- sort(table(waterNames), decreasing = TRUE)
waterDf <- data.frame(waterCounts)
waterDf <- waterDf %>% bind_rows(slice(., 11)) %>% slice(-11)
waterDf$waterNames <- factor(waterDf$waterNames, levels = waterDf$waterNames)

ggplot(waterDf, aes(waterNames, Freq)) +
  geom_bar(stat="identity") +
  geom_text(aes(label = Freq), nudge_y = 0.5) +
  labs(title = "Different Names for Water",
       x = "",
       y = "Counts") +
  scale_x_discrete(guide = guide_axis(angle = 30)) +
  theme_minimal() +
  theme(plot.title = element_text(size = 16, hjust = 0.5),
        axis.title.y = element_text(size = 14),
        axis.text = element_text(size = 12),
        panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank())
```

We will need to do a lot of data cleaning in order to analyze any of these ingredient lists. I will spare you most of the details. Briefly, I converted all of the ingredient names to lowercase, removed unnecessary characters, and merged many ingredients by hand (really I just used a bunch of regular expressions.)

Let's count how many unique ingredients we have after each cleaning step.

```{r echo=FALSE}
as.vector(do.call(c, ingredientLists, quote=T)) %>% unique %>% length %>% kable(col.names = "# of original ingredients", align = "l")

source(paste0(here(), "/src/ingredientRegex.R"))
ingredientListsLowerCase <- lapply(ingredientLists, tolower)
as.vector(do.call(c, ingredientListsLowerCase, quote=T)) %>% unique %>% length %>% kable(col.names = "# of lowercase ingredients", align = "l")

ingredientListsNoWeirdness <- lapply(ingredientListsLowerCase, str_replace_all, strangeCharactersRegex)
as.vector(do.call(c, ingredientListsNoWeirdness, quote=T)) %>% unique %>% length %>% kable(col.names = "# of ingredients w/o strange characters", align = "l")

names(ingredientRegex) <- tolower(names(ingredientRegex))
ingredientRegex <- tolower(ingredientRegex)
ingredientListsClean <- lapply(ingredientListsNoWeirdness, str_replace_all, ingredientRegex)
as.vector(do.call(c, ingredientListsClean, quote=T)) %>% unique %>% length %>% kable(col.names = "# of merged and cleaned ingredients", align = "l")
```

Now that we have cleaned our ingredient names and gone from 857 to 713 unique ingredients, we can do some actual analysis of the data.

### Number of ingredients

How many ingredients do each of these moisturizers contain?

```{r echo=FALSE}
ingredientsUnique <- as.vector(do.call(c, ingredientListsClean, quote=T)) %>% unique %>% sort

ingDf <- data.frame(row.names = ingredientsUnique)

for(i in productSheet$INCIDECODER_URL){
  ingDf[,i] <- as.numeric(match(row.names(ingDf), ingredientListsClean[[i]]))
}

ingDf[is.na(ingDf)] <- as.numeric(0)
ingDf[ingDf>0] <- as.numeric(1)

prodDf <- as.data.frame(t(ingDf))

ingredientCountsPerProduct <- data.frame(product = paste(productSheet$BRAND_NAME, productSheet$PRODUCT_NAME, sep = " "),
                                         count = rowSums(prodDf), 
                                         row.names = NULL)

ingredientCountsPerProductSummary <- data.frame(minimum = min(ingredientCountsPerProduct$count),
                                                median = median(ingredientCountsPerProduct$count),
                                                maximum = max(ingredientCountsPerProduct$count))

kable(ingredientCountsPerProductSummary)
```

Here are the top and bottom 10 products, sorted by number of ingredients.

```{r echo=FALSE, fig.width=10, fig.height=7}
ingredientCountsPerProductSorted <- arrange(ingredientCountsPerProduct, count)

numberOfProducts <- nrow(ingredientCountsPerProductSorted)
ingredientCountsTopAndBottom <- ingredientCountsPerProductSorted[c(1:10,                
                                                                   (numberOfProducts-10):numberOfProducts),]
ingredientCountsTopAndBottom$product <- factor(ingredientCountsTopAndBottom$product,
                                                      levels = ingredientCountsTopAndBottom$product)

ggplot(ingredientCountsTopAndBottom, aes(product, count)) +
  geom_bar(stat="identity") +
  labs(title = "Number of Ingredients in Moisturizers",
       x = "",
       y = "Counts") +
  scale_y_continuous(sec.axis = dup_axis()) +
  coord_flip() +
  theme_minimal() +
  theme(plot.title = element_text(size = 16, hjust = 0.5),
        axis.title.x = element_text(size = 14),
        axis.text = element_text(size = 12),
        panel.grid.minor = element_blank(),
        panel.grid.major.y = element_blank())

```

What is the deal with the Ole Henriksen moisturizer? Do they really need to use 72 ingredients? Does that make it a better moisturizer? 

### Price

We also have the price per ounce of each product. Here is a summary:

```{r echo=FALSE, fig.width=10, fig.height=7}
ingCountsPrice <- data.frame(ingredientCountsPerProduct,
                             priceperoz = productSheet$PRICE/productSheet$SIZE)

prodPriceSummary <- data.frame(minimum = min(ingCountsPrice$priceperoz),
                                     mean = mean(ingCountsPrice$priceperoz),
                                     median = median(ingCountsPrice$priceperoz),
                                     maximum = max(ingCountsPrice$priceperoz))

prodPriceSummary <- format(round(prodPriceSummary, 2), nsmall = 2)
prodPriceSummary <- paste0("$", prodPriceSummary)
prodPriceSummaryFormatted <- data.frame(minimum = prodPriceSummary[1],
                                     mean = prodPriceSummary[2],
                                     median = prodPriceSummary[3],
                                     maximum = prodPriceSummary[4])

kable(prodPriceSummaryFormatted)

ingredientPriceSorted <- arrange(ingCountsPrice, priceperoz)

ingredientPriceTopAndBottom <- ingredientPriceSorted[c(1:10, (numberOfProducts-10):numberOfProducts),]
ingredientPriceTopAndBottom$product <- factor(ingredientPriceTopAndBottom$product,
                                                      levels = ingredientPriceTopAndBottom$product)

priceLabels <- function(x) {
  paste0("$", x)
}

ggplot(ingredientPriceTopAndBottom, aes(product, priceperoz)) +
  geom_bar(stat="identity") +
  labs(title = "Price/Ounce",
       x = "",
       y = "") +
  scale_y_continuous(breaks = seq(0, 60, 10), labels = priceLabels, sec.axis = dup_axis()) +
  coord_flip() +
  theme_minimal() +
  theme(plot.title = element_text(size = 16, hjust = 0.5),
        axis.title.x = element_text(size = 14),
        axis.text = element_text(size = 12),
        panel.grid.minor = element_blank(),
        panel.grid.major.y = element_blank())
```


### Correlation of price and number of ingredients
Now, we can compare prices to the number of ingredients in each moisturizer. Let's do a quick correlation test first.

```{r echo=FALSE,include=TRUE,results='asis'}
ingredientCount <- ingCountsPrice$count
pricePerOz <- ingCountsPrice$priceperoz

cat('<pre>')
print(cor.test(ingredientCount, pricePerOz))
cat('</pre>')
```

There is a statistically significant correlation between price and ingredient count. I guess that makes sense. Let's visualize the correlation using the [plotly R package](https://plotly.com/r/) to give some interactivity.

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=7}
library(plotly)

ingCountsPrice$store <- productSheet$STORE

p <- ggplot(ingCountsPrice, aes(count, priceperoz)) +
  geom_point(aes(text = product, colour = store)) +
  geom_smooth() +
  scale_x_continuous(breaks = seq(0, 70, 10)) +
  scale_y_continuous(breaks = seq(0, 60, 10), labels = priceLabels) +
  scale_color_manual(values = c("target" = "#cc0100", "sephora" = "#143642")) +
  labs(title = "hover to see product names",
       x = "Number of Ingredients",
       y = "Price/Ounce") +
  theme_minimal() +
  theme(plot.title = element_text(size = 12, hjust = 0.5),
        axis.title = element_text(size = 14),
        axis.text = element_text(size = 12))

ggplotly(p, tooltip = "text") %>%
  style(hoverinfo = "skip", traces = 3) %>%
  style(hoverinfo = "skip", traces = 4)
```

Interestingly, it looks like the Ole Henriksen moisturizer isn't so bad if you're just looking to maximize the number of ingredients in your lotion. Bobbi Brown and Charlotte Tilbury, on the other hand, are out of control. How are people spending more than $50 per ounce of moisturizer? More dollars than sense, I guess. 

### Ingredient prevalence

What are the most and least common ingredients in our moisturizer dataset? Let's take a look at the distribution of how many products each ingredient is found in. 

```{r echo=FALSE}
ingCountDf <- data.frame(ingredient = row.names(ingDf),
                         count = rowSums(ingDf), 
                         row.names = NULL)

ingCountSummary <- data.frame(minimum = min(ingCountDf$count),
                              mean = mean(ingCountDf$count),
                              median = median(ingCountDf$count),
                              maximum = max(ingCountDf$count))

kable(ingCountSummary)
```

Ok, the median is 1. That means most of the ingredients are found in only one product. That makes sense because we have 713 ingredients and 74 products. Let's try to visualize this distribution with a histogram.

```{r echo=FALSE, warning=FALSE, fig.width=10, fig.height=7}
ggplot(ingCountDf, aes(count)) +
  geom_histogram(binwidth = 1) +
  labs(title = "Histogram of Ingredient Prevalence",
       x = "Number of products with each ingredient",
       y = "Count") +
  scale_x_continuous(breaks = seq(0, 75, 5)) +
  scale_y_continuous(breaks = seq(0, 400, 100)) +
  theme_minimal() +
  theme(plot.title = element_text(size = 16, hjust = 0.5),
        axis.title.x = element_text(size = 14),
        axis.title.y = element_text(size = 14),
        axis.text = element_text(size = 12),
        panel.grid.minor = element_blank())
```

Wow, the distribution is heavily skewed towards ingredients that appear in very few products. Here is a table, counting the ingredients with that are not present in many products:

```{r echo=FALSE, fig.width=2}
rareIngredientCounts <- NULL
for(i in 1:5){
  rareIngredientCounts <- c(rareIngredientCounts, nrow(ingCountDf[ingCountDf$count == i,]))
}

rareIngredientDf <- data.frame("count" = rareIngredientCounts, row.names = c("number of ingredients used in 1 product", paste("number of ingredients used in", 2:5, "products")))

kable(rareIngredientDf, align = 'r', format = "html", table.attr = "style='width:40%;'")
```
\
\
That's 413/713 ingredients that are only in 1 product, and 629/713 that are in less than 6 of the products. We do, however, have some ingredients that are used in most of the moisturizer products.

```{r echo=FALSE, warning=FALSE, fig.width=10, fig.height=7}

ingCountSorted <- arrange(ingCountDf, count)

ingCountTop <- ingCountSorted[(nrow(ingCountSorted)-20):nrow(ingCountSorted),]

ingCountTop$ingredient <- factor(ingCountTop$ingredient,
                                 levels = ingCountTop$ingredient)

ingCountTop$percent <- 100*ingCountTop$count/nrow(productSheet)

percentLabels <- function(x) {
  paste0(x, "%")
}

ggplot(ingCountTop, aes(ingredient, percent)) +
  geom_bar(stat="identity") +
  labs(title = "20 Most Common Ingredients",
       x = "",
       y = "Percentage of products") +
  scale_y_continuous(breaks = seq(0, 100, 20),
                     labels = percentLabels,
                     sec.axis = dup_axis()) +
  coord_flip() +
  theme_minimal() +
  theme(plot.title = element_text(size = 16, hjust = 0.5),
        axis.title.x = element_text(size = 14),
        axis.text = element_text(size = 12),
        panel.grid.minor = element_blank(),
        panel.grid.major.y = element_blank())
```

It turns out that water and glycerin are used in almost all of the moisturizers. 


```{r fig.width=20, fig.height=20, warning=FALSE, error=FALSE, message=FALSE, echo=FALSE}
library(here)
library(rvest)
library(gplots)
library(RColorBrewer)
library(Rtsne)
library(ggplot2)
library(ggrepel)
library(dplyr)

par(mar=c(1,1,1,1))
par(oma=c(1,1,1,1))
par(cex.main = 1.5)

#### read moisturizer data (generated with getMoisturizerData.R) ####
## TODO reformat column names then remove check.names=FALSE
# read and format data
moistData <- read.csv(paste0(here(), "/data/moisturizer_data.csv"), check.names = FALSE)

prods <- subset(moistData, select = -c(BRAND_NAME, PRODUCT_NAME, PRICE, SIZE, PRICE_PER_OZ, STORE))
prods$product <- paste(moistData$BRAND_NAME, moistData$PRODUCT_NAME)
# move product column to first column
prods <- relocate(prods, product)

# keep ingredients that are used by more than one product
ingCols <- prods[,-1]
ingShared <- ingCols[, which(colSums(ingCols) > 20)]
# prodsIngShared <- cbind(prods$product, ingShared)
# names(prodsIngShared)[1] <- "product"

row.names(ingShared) <- prods$product

prodDf <- t(ingShared)


M <-cor(prodDf, method = "spearman")
corrplot(M, type="upper", order="hclust",
         col=brewer.pal(n=8, name="RdYlBu"))


```